\plaintitle

### 2017-04-26 statistical-quorum.tex
Style section names.

@@

\begin{table*}
  \centering
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
  invalid votes. The last column shows computed decision credibility as described in Section \nameref{sec:statistical-quorum}. We see that
  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
  with more votes any option could end up winning.}\label{tab:example}
\end{table*}

To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
hypothesis testing to evaluate the credibility of the voting result.
It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
and not just those who voted.
Intuitively, if in a community of 100 members 40 members vote for a decision to pass and nobody against,
we can see that if everyone voted, the probability of majority of population voting for a decision is high.
On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
be if everyone voted.
Only two more votes against can completely change the outcome.
Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
We describe decision credibility in more detail in Section \nameref{sec:statistical-quorum}.

Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
has to pass for a decision to be seen as legitimate.
For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
which puts legitimacy of those decisions into a question.
We discuss a way for a community to determine a reasonable statistical quorum in Section \nameref{sec:discussion}.

@@

We built an open source web application PeerMind integrating statistical quorum with online voting.
The application exposes computed decision credibility to users in real-time and through that provides
the feedback loop needed for improving utilization of collective attention.
We present the application in Section \nameref{sec:peermind} and use it for evaluating statistical quorum in
Section \nameref{sec:evaluation}.

@@

% theoretical: math, diagrams

>>>>

\newcommand\secref[1]{\textit{\nameref{#1}}}

@@

\begin{table*}
  \centering
  \begin{tabular}{l r r r r r r r}
    {\small\textit{Vote}} & \multicolumn{2}{r}{\small \textit{First option}} & \multicolumn{2}{r}{\small \textit{Second option}} & {\small \textit{Total eligible voters}} & {\small \textit{Turnout}} & {\small \textit{Decision credibility}} \\
    \midrule
    Brexit & 17,410,742 & $51.9\%$ & 16,141,241 & $48.1\%$ & 46,500,001 & $72.2\%$ & $59.8\%$ \\
    Presidential elections & 62,984,825 & $48.9\%$ & 65,853,516 & $51.1\%$ & 230,585,915 & $55.9\%$ & $52.6\%$ \\
  \end{tabular}
  \caption{Votes cast in Brexit, a 2016 referendum for United Kingdom's withdrawal from the European Union~\protect\cite{brexitresults},
  and 2016 USA presidential elections, popular vote~\protect\cite{usaresults1,usaresults2}. Simplified to only two options and without
  invalid votes. The last column shows computed decision credibility as described in Section \secref{sec:statistical-quorum}. We see that
  credibility is low, which we can mostly attribute to the fact that there is small difference in votes between both options so
  with more votes any option could end up winning.}\label{tab:example}
\end{table*}

To answer this question we are proposing a novel metric, \emph{decision credibility}, which uses statistical
hypothesis testing to evaluate the credibility of the voting result.
It uses distribution of votes made to evaluate the probability of decision passing if everyone actually voted,
and not just those who voted.
Intuitively, if in a community of 100 members 40 members vote for a decision to pass and nobody against,
we can see that if everyone voted, the probability of majority of population voting for a decision is high.
On the other hand, if 41 members for in favor and 40 against, we cannot really say much about what would a decision
be if everyone voted.
Only two more votes against can completely change the outcome.
Hence, the suggested action would be for the community to gather more votes and continue discussion around the decision.
We describe decision credibility in more detail in Section \secref{sec:statistical-quorum}.

Based on decision credibility we can define a \emph{statistical quorum} as a threshold which a decision's credibility
has to pass for a decision to be seen as legitimate.
For example, if we adopt $90\%$ as a reasonable requirement for statistical quorum and we compute decision
credibility for votes in Table~\ref{tab:example} we see that in both cases we do not reach statistical quorum
which puts legitimacy of those decisions into a question.
We discuss a way for a community to determine a reasonable statistical quorum in Section \secref{sec:discussion}.

@@

We built an open source web application PeerMind integrating statistical quorum with online voting.
The application exposes computed decision credibility to users in real-time and through that provides
the feedback loop needed for improving utilization of collective attention.
We present the application in Section \secref{sec:peermind} and use it for evaluating statistical quorum in
Section \secref{sec:evaluation}.

@@
